---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, eval = T, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chatr

<!-- badges: start -->
<!-- badges: end -->

The goal of chatr is to allow researchers to easily set up Chatter experiments.

To get started, visit [Chatter](http://chatter-washu.herokuapp.com/research/sign_in) and sign up for a researcher account.  You'll receive an email when your account is approved.

Log in and find your API credentials.  You'll need the Token for any calls you make to Chatter with `chatr`.  Set it as an environment variable.

```{r, eval=F}
Sys.setenv(BEARER_TOKEN = )
```


## Installation

You can install the `chatr` package from Github.

```{r}
#install.packages("devtools")
devtools::install_github("erossiter/chatr",auth_token = )
```

After install, you should load the package.

```{r, eval=T}
library(chatr)
```


## Example

Running an experiment with `chatr` requires seven basic steps.  I'll walk through a basic example here.

1. Create an experiment
2. Create instructions
3. Create chatrooms
4. Create users
5. Create chatroom memberships
6. Run the experiment
7. Download results

### Create an experiment

First, you create an experiment with the `create_experiment` function.  All that is required is a name for the experiment.

```{r eval=T, echo=T}
my_experiment <- create_experiment(name = "My First Experiment")
my_experiment
```
We'll note the experiment id.  We'll need it moving forward.

```{r}
(experiment_id <- my_experiment$content$id)
```

### Create instructions

Next, we create instructions for our users.  For this experiment, I have two unique instructions.  Users will be assigned to either the treatment or control instructions.  Note that we associate these instructions with the experiment we just created via the `experiment_id` argument  The `text` argument is what users see.

```{r}
instruction_t <- create_instruction(experiment_id = experiment_id,
                   name = "Treatment",
                   text = "Please talk about X.")

instruction_c <- create_instruction(experiment_id = experiment_id,
                   name = "Control",
                   text = "Please talk about Y.")
```

We'll note the instruction ids as well.

```{r}
(treatment_id <- instruction_t$content$id)
(control_id <- instruction_c$content$id)
```



### Create chatrooms

Next, let's create ten chatrooms.  The `topic` argument is the unique name for the chatroom.  Note that users assigned to each chatroom (a step we'll complete momentarily) can see this name. The `min_duration` argument indicates how long participants must stay in the chatroom in seconds.

```{r}
for(i in 1:10){
  create_chatroom(topic = paste0("chatroom", i),
                min_duration = 180)
}
```

We can see that our chatrooms have been created with another `chatr` function: `list_chatrooms`.  This function returns a list, and we're interested in the `content` element right now holding a dataframe with information about our chatrooms.

```{r}
my_chatrooms <- list_chatrooms()
my_chatrooms$content
```
We'll note the chatroom ids.

```{r}
(chatroom_id <- my_chatrooms$content$id)
```


### Create users

Next we'll create 20 users and check them out. The `username` argument is the unique name given to each user.  This is the name a user would use to log in to Chatter^[Logging in is not required with the use of deeplinks, explained here.] and is the default display name^[Display names can be customized when assigning chatroom members, explained here.] that others in the user's chatroom see.  

```{r}
for(i in 1:20){
  create_user(username = paste0("user", i))
}
```

```{r}
my_users <- list_users()
head(my_users$content)
```
We'll again note the ids.

```{r}
(user_id <- my_users$content$id)
```

### Create chatroom memberships

Finally, we need to tie it all together.  This is where the magic happens!  We need to create chatroom memberships, and we do so by assigning instructions and chatroom to users.  To help visualize what we're doing ahead of time, let's create a dataframe where each row is a user, and we note which chatroom they're assigned to and which instruction they should see.


```{r}
# each row is a user
exp <- data.frame(user_id = user_id)

# consecutive users are put in a chatroom together
exp$chatroom_id <- rep(chatroom_id, each = 2)

# the first five chatrooms are in the treatment,
# the last five are in the control
exp$instruction_id <- rep(c(treatment_id, control_id), each = 10)
```

Finally, we tell Chatter with the `create_chatroom_memberhip` function.  The required arguments are the IDs for the user, the instruction they should see, and the chatroom they should be in. 

```{r}
for(i in 1:nrow(exp)){
  create_chatroom_membership(user_id = exp$user_id[i],
                             chatroom_id = exp$chatroom_id[i],
                             instruction_id = exp$instruction_id[i])
}
```

```{r, eval=T, echo=F}
my_chatroom_memberships <- list_chatroom_memberships()
```

At this point, you should be able to [log in to Chatter](http://chatter-washu.herokuapp.com/login) as any of the users you created.

### Other functionality

`chatr` has the following additional functionality.  See package documentation for more information.  

- `create` create an entry
- `update` update field(s) for an entry
- `retrieve` retrieve an entry
- `list` list all entries
- `delete` delete an entry (you can delete this example with the following code, note the order of deletion is opposite the order of creation due to the dependencies of)

```{r, echo=T, eval=T, results = 'hide'}
for (i in 1:length(my_chatroom_memberships$content$id)) {
  delete_chatroom_membership(my_chatroom_memberships$content$id[i])
}

for (i in 1:length(chatroom_id)) {
  delete_chatroom(chatroom_id[i])
}

for (i in 1:length(user_id)) {
  delete_user(user_id[i])
}

delete_instruction(treatment_id)
delete_instruction(control_id)
delete_experiment(experiment_id)
```



## Full Example with `pyMTurkR` and `chatr`

Chatter experiments can be implemented with Amazon Mechanical Turk participants.  Here I'll walk though a short example using the [`pyMTurkR`](https://github.com/cloudyr/pyMTurkR) package to make calls to the MTurk API.  First, follow all of the instructions on the [`pyMTurkR` Github repo](https://github.com/cloudyr/pyMTurkR) to install it.

I recommend playing in the MTurk Requester Sandbox.  Recall, to test you're in the Sandbox, your account balance should be $10,000.00

```{r}
pyMTurkR::AccountBalance()
```

### Recruitment HIT

One way to conduct a Chatter experiment is to recruit participants, say in the morning, and invite them back, say in the afternoon, for the experiment.  On MTurk, one way to do this is by using two HITs.

Here's an example of an initial recruitment HIT.  I first create three qualifications for the HIT---participants must have a HIT approval rating about 95%, they must have completed 1000 HITS, and they must be verified as living in the US.  

```{r}
qual_approvehits <- list(QualificationTypeId = '000000000000000000L0',
                         Comparator = 'GreaterThanOrEqualTo',
                         IntegerValues = 95,
                         ActionsGuarded = "DiscoverPreviewAndAccept")
qual_numhits <- list(QualificationTypeId = '00000000000000000040',
                     Comparator = 'GreaterThanOrEqualTo',
                     IntegerValues = 1000,
                     ActionsGuarded = "DiscoverPreviewAndAccept")
qual_usa <- list(QualificationTypeId = '00000000000000000071',
                 Comparator = 'EqualTo',
                 LocaleValues = list(Country = 'US'),
                 ActionsGuarded = "DiscoverPreviewAndAccept")
hit1_req <- pyMTurkR::GenerateQualificationRequirement(quals = list(qual_approvehits,
                                                                    qual_numhits,
                                                                    qual_usa))
```

Second, I load in the "question" associated with the HIT.  This is an HTML formatted page where I include a Qualtrics URL to direct participants to the researcher's pre-treatment survey.

```{r}
f <- system.file("templates/instructions_treatment_rep.html", package = "chatr")
hit1_question <- pyMTurkR::GenerateHTMLQuestion(file = f)
```

Finally, the HIT itself gives basic information about the task so MTurkers know if they want to proceed or not, in addition to the requirements and question created above.

```{r}
pyMTurkR::CreateHIT(assignments = 100,
                    expiration = 240, 
                    duration = 900,
                    reward = '0.75',
                    title = 'Demographics & politics survey (3-5min).
                    Qualify for above min wage HIT! Take now, expires soon!',
                    keywords = 'survey, demographics, politics',
                    description = "Share political opinions on salient issues.
                    Answer demographic questions.  May qualify you for follow up
                    HIT ($2, $1 bonus) that will be live at the top of the hour.",
                    question = hit1_question,
                    qual.req = hit1_req)
```



### Prep experiment

Assume the recruitment HIT was a success and 100 participants took the survey.  Before bringing participants back to the Chatter experiment, the researcher needs to group the participants for conversations and randomly assign any treatments.  Here, I'll demonstrate a simple pairing algorithm with simulated data included in the `chatr` package.  Then, given the resulting partnerships and treatment assignments from the algorithm, I'll demonstrate how to upload this information to Chatter with `chatr` functions.  

#### Simple pairing algorithm

Load the example recruitment HIT data from the `chatr` package, including a fake ID variable and a binary indicator of partisanship. (Presumably a researcher's pre-treament survey would also ask many covariates, excluded for simplicity.)

```{r}
data("recruit_ex") #load example data
head(recruit_ex)
```

The following pairing algorithm is simple.  For each participant, check to see if they already have a partner.  If so, skip them.  If not, randomly choose their partner from those who are of the opposite partisanship and aren't already in a partnership.  Assign the new partnership a unique ID.  Iterate through participants until as many partnerships as possible are made.

```{r}
set.seed(522)
recruit_ex$pair_id <- NA
pair_id_iter <- 1
for (i in 1:nrow(recruit_ex)) {
  if (!is.na(recruit_ex$pair_id[i])) {
    next
  }
  
  pid <- recruit_ex$pid[i]
  poss_partners <- which(is.na(recruit_ex$pair_id) & (recruit_ex$pid != pid))
  if (length(poss_partners) == 0) {
    next
  } else if (length(poss_partners) == 1) {
    partner_idx <- poss_partners
  } else {
    partner_idx <- sample(poss_partners, size = 1)
  }
  
  recruit_ex$pair_id[i] <- pair_id_iter
  recruit_ex$pair_id[partner_idx] <- pair_id_iter
  pair_id_iter <- pair_id_iter + 1
}
```

The next step is to assign treatment at the partnership (or cluster) level.  In this experiment, partnerships can either be assigned to conversation as a form of contact with their partner or no contact with their partner.

```{r}
recruit_ex$Z <- randomizr::cluster_ra(recruit_ex$pair_id)
```

We can see the pairing algorithm and treatment assignment worked.  Partnerships each feature one Democrat and one Republican.  We can also see the treatment status (`Z`) of the partnership.

```{r}
head(recruit_ex[order(recruit_ex$pair_id),])
```
#### chatr

`chatr` makes it easy to create our experiment with the Chatter software.  First, use the `create_experiment` function to create a new experiment.  The `post_survey_link` is any post-treatment Qualtrics survey the researcher wants participants to take.  Participants will be redirected there after their time using Chatter.

```{r}
my_experiment <- create_experiment(name = "Partner contact experiment",
                                   post_survey_link = "http://erossiter.com/")
```

Next, we need to create instructions for our participants.  In this experiment, we'll have three unique instructions.  One for control, one for treated Republicans (telling them they are talking to a Democrat), and one for treated Democrats (telling them they are talking to a Republican).  

Instructions can be formatted with HTML.  Here, I load example instructions from the `chatr` package. 


```{r}
f_control <- system.file("templates/instructions_control.html", package = "chatr")
f_treatment_d <- system.file("templates/instructions_treatment_d.html", package = "chatr")
f_treatment_r <- system.file("templates/instructions_treatment_r.html", package = "chatr")

html_control <- paste(readLines(f_control), collapse="")
html_treatment_d <- paste(readLines(f_treatment_d), collapse="")
html_treatment_r <- paste(readLines(f_treatment_r), collapse="")

instruction_c <- create_instruction(experiment_id = my_experiment$content$id,
                   name = "Control",
                   text = html_control)

instruction_t_d <- create_instruction(experiment_id =  my_experiment$content$id,
                   name = "Treatment for Dems",
                   text = html_treatment_d)

instruction_t_r <- create_instruction(experiment_id =  my_experiment$content$id,
                   name = "Treatment for Reps",
                   text = html_treatment_r)
```

For convenience, let's now add which instruction we intend for each participant to have to our `recruit_ex` dataframe.  We'll assign this instruction to the user via `chatr` below.

```{r}
recruit_ex$instruction_id <- NA
recruit_ex$instruction_id[recruit_ex$Z == 1 & recruit_ex$pid == "D"] <- instruction_t_d$content$id
recruit_ex$instruction_id[recruit_ex$Z == 1 & recruit_ex$pid == "R"] <- instruction_t_r$content$id
recruit_ex$instruction_id[recruit_ex$Z == 0] <- instruction_c$content$id
```

Next, we'll create 50 chatrooms and 100 users for our Chatter experiment.  As we create them, we'll add which room and user we intend to be associated with each participant to our `recruit_ex` dataframe.

```{r}
for(i in unique(recruit_ex$pair_id)){
  room <- create_chatroom(topic = paste0("chatroom_pair", i),
                min_duration = 480)
  recruit_ex$chatroom_id[recruit_ex$pair_id == i] <- room$content$id
}

for(i in 1:nrow(recruit_ex)){
  user <- create_user(username = paste0("user", recruit_ex$id[i]))
  recruit_ex$user_id[i] <- user$content$id
}
```

So far, we've created instructions, chatrooms, and users separately in Chatter.  Now, we need to tie things together by creating chatroom memberships that tell Chatter what instruction and chatroom should be associated with each user.  Before doing so, let's give each user an optional display name that others in the chatroom see alongside the user's messages.

```{r}
recruit_ex$display_name <- stringi::stri_rand_strings(nrow(recruit_ex), 5)

for(i in 1:nrow(recruit_ex)){
  create_chatroom_membership(user_id = recruit_ex$user_id[i],
                           chatroom_id = recruit_ex$chatroom_id[i],
                           instruction_id = recruit_ex$instruction_id[i],
                           display_name = recruit_ex$display_name[i])
}
```

That's it!  Chatter is ready for participants to log in.

### An aside on deeplinks

[Deep linking](https://en.wikipedia.org/wiki/Deep_linking) is useful with Chatter experiments.  Deeplinks allow the researcher to provide a participant with a URL to Chatter that logs the participant in to view their assigned chatroom(s).  This obviates the need for participants to have a username for Chatter and log themselves in.


Chatter deeplinks take the following form:

`http://chatter-washu.herokuapp.com/deeplink?user=<username>&room=<chatroom_id>`

The `room` field is optional. Specifying the `room` field will bring a user *into* that specific chatroom.  Without the `room` field specified, the user will be logged into Chatter and see a list of chatrooms they are assigned to.

If you didn't delete the example experiment created above, you can use the following deeplinks to see the difference.

- `http://chatter-washu.herokuapp.com/deeplink?user=user1&room=2631`
- `http://chatter-washu.herokuapp.com/deeplink?user=user1`



### Experiment HIT

So far, we've recruited participants for our study and prepped our Chatter experiment.  Now, we need to invite participants back and ask them to complete our second HIT.  Before posting the second HIT for the experiment, researchers will want to create and assign a qualification for participants invited back.  

```{r, eval=F}
chatter_qual <- pyMTurkR::CreateQualificationType(name = "Invited",
                                   description = "Invited to Chatter experiment",
                                   status = "Active")
```

```{r eval=F}
pyMTurkR::AssignQualification(qual = chatter_qual$QualificationTypeId,
                              workers = recruit_ex$id, #recall these are fake ids
                              value = 100,
                              notify = T)
```

And, researchers will want to require that the Workers have a score of 100 on the qualification we just created to preview and accept the HIT.

```{r, eval = F}
qual_chat <- list(QualificationTypeId = chatter_qual$QualificationTypeId,
                 Comparator = 'EqualTo',
                 IntegerValues = 100,
                 ActionsGuarded = "DiscoverPreviewAndAccept")
hit2_req <- pyMTurkR::GenerateQualificationRequirement(quals = list(qual_chat))
```

I've also found it useful to send a reminder email to MTurk Workers before the HIT is live.

```{r eval=F}
pyMTurkR::ContactWorker(subjects = "Reminder for follow-up HIT",
                        msgs = "Please return for follow-up HIT",
                        workers = recruit_ex$id) #recall these are fake ids
```

As with the recruitment HIT, we need to load our question HTML for the experiment HIT.  A template is stored in the `chatr` package.^[I want to caution that I've never used this template in my research. Please test thoroughly for any bugs. See [this resource](https://research-it.wharton.upenn.edu/news/capture-mturk-workers-ids-qualtrics-survey/).]  An important feature of this template is its use of deeplinks.  The HTML includes a deeplink created given the user's MTurk id, collected as a variable.  The usernames in our experiment are prefaced with "user", so the specific deeplinks in the `templates/chatter_question.xml` file take the form of: `http://chatter-washu.herokuapp.com/deeplink?user=user` + `mturkIDvariable`

```{r, eval = F}
f <- system.file("templates/test.xml", package = "chatr")
hit2_question <- pyMTurkR::GenerateHTMLQuestion(file = f)
```

Finally, we post the HIT!  

```{r, eval=F}
pyMTurkR::CreateHIT(assignments = 100,
                    expiration = 600, 
                    duration = 3600,
                    reward = '2',
                    title = 'Follow up HIT: Conversation/short essay',
                    keywords = 'survey, conversation, short essay, follow up',
                    description = "This is a follow up HIT for Workers
                    that completed initial survey on demographics and political
                    opinions.  Potential $1.00 bonus.",
                    question = hit2_question,
                    qual.req = hit2_req)
```

All Chatter data can be viewed on the Researcher GUI, or can be downloaded with the `list` functions in the `chatr` package.


```{r, echo=F, eval=T, results = 'hide'}
# Wipe my account
cont <- T
while (cont) {
  m <- list_messages()$content$id
  cm <- list_chatroom_memberships()$content$id
  c <- list_chatrooms()$content$id
  u <- list_users()$content$id
  ins <- list_instructions()$content$id
  e <- list_experiments()$content$id

  if (length(c(u, c, cm, ins, e, m)) == 0) cont <- F
  
  for (i in 1:length(m)) try(delete_message(m[i]))
  for (i in 1:length(cm)) try(delete_chatroom_membership(cm[i]))
  for (i in 1:length(c)) try(delete_chatroom(c[i]))
  for (i in 1:length(u)) try(delete_user(u[i]))
  for (i in 1:length(ins)) try(delete_instruction(ins[i]))
  for (i in 1:length(e)) try(delete_experiment(e[i]))
}
```


## More examples of Chatter experiments in practice

There are many ways to use Chatter in practice.

- You could collect pre-treatment information and set up chatrooms amongst specific participants.  Here, you could email participants a specific link to join, use deeplinks as demonstrated with the MTurk example above, etc.
- You could have participants create their own Chatter account (and thus pick their own username), and use the API to assign them to chatrooms (with a script in real time or not.)
- You could reprogram an experiment and feed participants into the chatrooms in real time as they are taking your survey.
- etc.

Broadly, these options fall into two camps.

1. *Coordinated chats* Participants are assigned to pre-specified chatrooms
2. *Real-time chats* Participants are assigned to chatrooms by convenience upon completing the pre-treatment survey.

I've demonstrated option #1 above.  I'll give an example of how to implement option #2 with the Qualtrics survey platform here.

### Real-time chats

Real-time chats can be created with deeplinks, too.  Again, you need to create an experiment in Chatter just as we did before.  The difference is that you will not know ahead of time exactly which participant will be filtered into each chatroom.

In Qualtrics, the idea is to increment an embedded data variable each time someone completes the survey.  If I'm the first participant to finish the survey, I'm "user1" and it is already programmed that I am filtered into "chatroom1".  The next person completes the survey, they are "user2", and join me in "chatroom1". 

TODO: When I get my Qualtrics account back I have an example of this. https://www.qualtrics.com/support/survey-platform/survey-module/survey-tools/quotas/#UsingQuotasInOtherPartsOfQualtrics

`http://chatter-washu.herokuapp.com/deeplink?user=EMBEDDED_DATA`

TODO: A more advance feature is real-time chats with a script prepared to make chatrooms on the fly...
